---
title: "WQP Overlaps"
author: "Matthew Ross"
date: "1/3/2018"
output: html_document
---

# Join WQP inventory data to the landsat wrs tiles. 

In order to avoid sending up extraneous data to the WQP portal, we check for the PATH and ROW of each site that is near landsat-detectable water (from step "1_landsat_site_inventory.Rmd"). Eventually this data will be joined to the actuall concentration data in 3_flat_overpasses.Rmd

```{r setup}
library(knitr)
library(feather)
library(googledrive)
library(tidyverse)
library(sf)
library(USAboundaries)

#Reset working directory to the top level
opts_knit$set(root.dir='../..')
```

## Load in site data and convert to an sf object
```{r}
#Load in wqp data
winv <- read_feather(sc_retrieve('2_rsdata/out/unique_site_visible_inv.feather.ind'))

#Convert the data into sf object
inv.sf <- st_as_sf(winv,coords=c('long','lat'),crs=4326) %>%
  st_transform(.,2163)

```


```{r}

#Read in usa data
usa = us_states() %>%
  st_transform(.,2163)


#Download and and read in WRS spatial data
download.file('https://landsat.usgs.gov/sites/default/files/documents/wrs2_asc_desc.zip', destfile = paste0(getwd(),'/2_rsdata/in/wrs2_asc_desc.zip'), method = 'curl')

unzip('2_rsdata/in/wrs2_asc_desc.zip', exdir = '2_rsdata/in/')

wrs <- st_read('2_rsdata/in/wrs2_asc_desc/wrs2_asc_desc.shp') %>%
  st_transform(.,2163)


#Negative buffer on the wrs tiles in usa by 1 km
wrs.usa <- wrs %>%
  st_buffer(.,0) %>%
  st_intersection(.,usa) %>%
  group_by(PATH,ROW) %>%
  st_buffer(.,-1000) %>%
  na.omit(.) %>%
  ungroup(.) 


#The above operation clips the wrs tiles to the shape of the usa, which we don't want. So let's keep full tiles
wrs.all.usa <- wrs %>%
  filter(PR %in% wrs.usa$PR) %>%
  select(PR_ID,PATH,ROW) %>%
  st_buffer(.,-1000) %>%
  mutate(area.km2=st_area(.)/(1000*1000)) %>%
  arrange(desc(area.km2)) %>%
  #Most of the tile areas are less than 32000 km2
  #But there are 3 with huge areas that don't have any data in them
  #Lets drop those %>% 
  mutate(pr=paste(PATH,ROW,sep='-')) %>%
  filter(!pr %in% c('23-153','133-91','24-153')) %>%
  select(-pr)



```

#Join inventory to wrs tiles
```{r}
#Use st join to join inventory to wrs path rows
inv.wrs.sf <- st_join(inv.sf,wrs.all.usa) %>%
  st_transform(.,4326)

#Convert this to a tibble
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- do.call(rbind,sf::st_geometry(x))
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

inv.wrs <- sfc_as_cols(inv.wrs.sf,names=c('long','lat')) %>%
  as.tibble() %>%
  dplyr::select(-geometry,-PR_ID) 

#Save data
data_file <- '2_rsdata/out/site_inventory_path_row.feather'
write_feather(inv.wrs,path=data_file)

#Upload to team drive
gd_put(scipiper::as_ind_file(data_file), data_file)
```




